Modeling How Evolution in One Trait is Affected by Adaptation in Another

### 2017-09-30 manuscript.tex
more updates to manuscript and graphs

\graphicspath{{figures/}}

@@

For quantitative traits, genetic associations and their effects on adaptation are mathematically described by Lande's equation ($\Delta \bar{z} = \textbf{G} \hspace{.05in}\beta$). It states that the change in the population mean $\Delta \bar{z}$ equals the product of the selection gradient $\beta$ and what is known as the \textbf{G}-matrix, whose components are the additive genetic covariances between traits. Lande’s result has been used to extend much of our understanding of how phenotypes evolve under selection, yielding a rich set of ideas that emphasize the role of the \textbf{G}-matrix in directing adaptive evolution at both micro- and macro-evolutionary scales.\par
% * <masel@email.arizona.edu> 2017-07-29T12:43:16.916Z:
% 
% > When combined with the notions of adaptive landscapes
% The link to adaptive landscapes is not obvious, including to me. G-matrix theory is all continuous, and the only possible landscape is a phenotype-fitness map. The adaptive landscape you are going to deal with is a genotype-fitness map, ie a completely different object.
% 
% ^ <kgomez81@math.arizona.edu> 2017-08-07T14:10:30.176Z.

The behavior of the \textbf{G}-matrix over time largely depends on what are the causes of the genetic correlations. Correlations between quantitative traits can arise for a number of reasons, but they often have a genetic basis \citep{Saltz2017}. They arise from either pleiotropy, linkage disequilibria, or are the product of trans acting elements; the first two that are thought to be the most significant factors. Historically, there have been two schools of thought what the principle causes of genetic correlations are. The first was put forth by the Edinburgh school, which held that pleiotropy was principle source, since linkage disequilibrium could rapidly decay by recombination \citep[Chapter~20]{fox2006evolutionary}. This belief was largely motivated by interests the relationships between parent-offspring phenotypes within randomly mating populations, and its implications for evolution. The alternative view came out of the Birmingham school, which largely experimented with inbred lines of plants. Geneticists in this group believed that linkage disequilibria was much more important; a fact that often reflected in their experiments. Presently, the Edinburgh view is more widely accepted, and as a result, most models for the evolution of the \textbf{G}-matrix assume genetic correlations are pleiotropic and proceed with examining how evolutionary processes influence them. We know far less about the about the behavior genetic correlations from linkage disequilibria over time, but this is not much of an issue for most complex organisms.\par
% * <masel@email.arizona.edu> 2017-08-22T16:50:23.214Z:
% 
% >  most models for the evolution of the \textbf{G}-matrix assume genetic correlations are pleiotropic and proceed with examining how evolutionary processes influence them
% An important exception to note is the "Bulmer effect", which I believe describes how selection can cause LD even in the absence of linkage, which then of course appears in G. This is distinct from our work where linkage to the background on which a mutation first appear is the cause of LD, albeit later shaped by selection. I believe there is some material on the Bulmer effect in Bruce Walsh's book. It is important to review it just enough to distinguish what you are doing from what Bulmer already did.
% 
% ^.
% * <masel@email.arizona.edu> 2017-08-22T16:47:43.713Z:
% 
% > the product of trans acting elements
% This is not self-explanatory, and insofar as I can guess what it means, it seems to me to be a special case of pleiotropy rather than a third distinct possibility.
% 
% ^.

@@

Lande’s equation describes the phenotypic evolution of populations with multiple characters from generation to generation.  Whether one can use this framework to uncover the history of natural selection in populations over large evolutionary timescales largely depends on the forces that influence the evolution of the G-matrix.  Over the past few decades there has been a significant amount of effort to shed light on the matter of G-matrix stability.  The analysis of genetic models for the adaptive evolution of multiple phenotypic traits indicate some of the conditions needed for stability\citep{Turelli1988,Jones2003}.  These works exclude the potential influences of linkage disequilibrium on the dynamics of variances and covariances over time, and as a result, our understanding of the role of linkage disequilibria in G-matrix evolution remains unclear.  Nonetheless, the question is significant there is empirical evidence suggesting that in the G-matrix is often conserved in natural populations \citep{Arnold1999,Roff2000,Steppan2002}.  However, much of what we know about G-matrix stability corresponds the examining evolutionary forces on

@@

% *****************************************************************************************************************
% ***********************************FIGURE 1*2D*Travelling*Wave***************************************************
% *****************************************************************************************************************
\begin{figure}
\includegraphics[width=1\linewidth]{fig1a.pdf}
\label{Figure 1.}
\caption{\footnotesize Evolutuion of a two dimensional traveling wave. The distribution depicted was sampled from simulations.  Each box represents a class composed of individuals with an identical number of beneficial mutations in each trait. Classes with solid coloring form part of the bulk, while those with patterned interiors represent the stochastic front. The red dot marks the mean number of beneficial mutations in each trait. The dashed red line represents a lines of constant fitness. Classes whose centers lie along this line have relative fitness equal to the population mean, but vary in the number of beneficial mutations that have accumulated in each trait. Relative fitness increases along the perpendicular direction. Classes . Beneficial mutations and random sampling produce negligible effects in the changes of their abundances. In asexuals, only beneficial mutations that occur on the fittest backgrounds produce lineages that may eventually fix. These are marked with arrows into the stochastic front. Other lineages are out-competed and remain insignificant in size with respect to bulk. As new classes, the bulk expands in the direction of increasing relative fitness, and classes below the red line shrink and eventually disappear. As a result, the two dimensional trait distribution continually shifts in the direction of increasing relative fitness over time. (Simulation parameters: $N=10^9$, $s=10^{-2}$, and $U=10^{-5}$)}
\end{figure}
% *****************************************************************************************************************
% *******************************END*FIGURE 1**********************************************************************
% *****************************************************************************************************************

The dynamics of the bulk was modeled by a system of ordinary differential equations, in which abundances changed in proportion to their selective advantage (\textbf{eqt. 1}). 
\begin{equation} \label{eq:1}
\dot{n}_{ij}(t) = (r_{ij}-\bar{r}) n_{ij}(t). 
\end{equation}
Abundances in the stochastic front were negligible in size relative to $N$ due to our parameter assumptions stated earlier.  Since the bulk represents the majority of population, we assumed that the frequencies of the bulk followed an analogous ODE.
\begin{equation} \label{eq:2}
\dot{p}_{ij}(t) = (r_{ij}-\bar{r}) p_{ij}(t) 
\end{equation}
These frequencies determine all of the quantitative genetic measures that are of interest to us.\par

@@

In order to investigate how of traits affect one another, we identified the contributions of each trait to the evolution of the population. The rate of adaptation can obtained by differentiating the population's mean relative fitness $\bar{r}=\sum_{ij} p_{ij} r_{ij}$, which is itself the sum of the two marginal means, $\bar{r_1}=\sum_{ij} p_{ij} is_1$ and $\bar{r_2}=\sum_{ij} p_{ij} j s_2$.  It follows that the total rate of adaptation is sum of rates of adaptation corresponding to each trait ($\dot{\bar{r}}=\dot{\bar{r_1}} +\dot{\bar{r_2}}$).  We can apply Lande's result to the two components $\dot{\bar{r_1}} $ and $\dot{\bar{r_2}} $ and rewrite them in terms of the trait variances $\sigma_1^2 = \sum_{ij} p_{ij} (r_i-\bar{r_1})^2$ and $\sigma_2^2 = \sum_{ij} p_{ij} (r_j-\bar{r_2})^2$, and their covariance $\sigma_{12} =\sum_{ij} p_{ij} (r_i-\bar{r_1})(r_j-\bar{r_2})$.  In doing so, we obtain the system expressed in equation 6.

@@

% Instead of hard-coding equation numbers, you can create auto-numbered links - ask Jason how. Right now the number doesn't match. You should also link this more clearly to the way it was presented in the introduction, ie you should explain that this is simply $\Delta \bar{z} = \textbf{G} \hspace{.05in}\beta$ in the case of two traits both contributing directly to fitness

@@

The contributions of the stochastic front were incorporated into the adaptive evolution of the population by simulating each establishment event. We did this by identifying the first mutant lineage destined to establish among all mutants appearing in the stochastic front. Mutants awhich are parallel to the line shownwhich are parallel to the line shownppear according to a time-inhomogeneous Poisson process with rate functions given by the abundances of the adjacent classes from which they originated. Consequently, the arrival time $\tau_l$ of the $l^{th}$ mutant generated from the class $n_{ij}(t)$ having a mutation in trait $k$ can be used to form the random variable
% * <masel@email.arizona.edu> 2017-08-22T22:55:54.046Z:
% 
% > can be used to form the random variable
% This is confusing. The equation is for the cumulant for the inhomogeneous rate, the equation does not describe a random variable.
% 
% ^.
\[ 
\int_0^{\tau_l} U_k n_{ij}(t) dt ,
\] 
which has the distribution of $l$ i.i.d exponentially distributed summed random variables. By randomly sampling from the this distribution, we could then solve for each arrival time $\tau_l$.   We then computed the probability of fixation using \textbf{equation 5}.  The integral was numerically approximated up to an appropriate cutoff.  To determine if a mutation would establish, we sampled a uniform random variable and checked it against the calculated $\pi_{fix}$. The next class to establish was chosen accordingly and incorporated into the bulk with a random initial size generated with from \textbf{equation 4}.\par
% * <masel@email.arizona.edu> 2017-08-22T22:57:52.494Z:
% 
% >  \textbf{equation 5}
% Again, replace all these with auto-numbered links to specific equations, not to hard-numbered text.
% 
% ^.

@@

We allowed the simulation to run for 5,000 generations to allow the population to achieve beneficial mutation-selection balance before collecting data. This ensured that the transient effects from the initializing would not distort the statistics collected for the distribution. Following the burn period of the simulation, we recorded the existing classes in the bulk and their respective abundances. The data was used to compute the marginal mean fitness and variances of each trait, as well as their covariance. From these quantities we also computed the instantaneous rate of adaptation (\textbf{eqtn. 5}), the mean rate of fitness increase in each trait and the overall rate of adaptation.

@@

In each of our simulations we observed a 40\% decrease in the rate of adaptation of a trait as a result adding a second. The decrease was consistent across three parameter sets which varied either $N$, $s$ and $U$. These results also agree with predictions that can be derived from Desai and Fisher's theoretical work ($v(N,s,U)$). However, our simulations show that upon adding the second trait, the net change in $v_1$ arises from an increase in the first trait's variance and the formation of negative covariance. The first results in an increase in $v_1$, as a direct consequence of equation 5. This gain is more than offset by the negative covariance created between the two traits, whose magnitude is always substantially larger than the variance increase.  Figure 2 presents the calculated scaled mean variance of trait one, and its scaled covariance with the second trait, as a function of each varying population parameter. The graph demonstrates that within these parameter ranges, the new rate of adaptation $v_1$ as a percentage of the prior in every case is nearly 60\% less due solely to the large amounts of negative covariance that builds up between the two traits.\par

% *****************************************************************************************************************
% ***********************************FIGURE 2*Scaled Var/Cov Changes & Trait Reduction*****************************
% *****************************************************************************************************************
\begin{figure}[h]
\begin{subfigure}[b]{1\linewidth}
\centering
\includegraphics[width=.48\linewidth]{fig5d.pdf}
\includegraphics[width=.48\linewidth]{fig5e.pdf}
\end{subfigure}
\begin{subfigure}[b]{1\linewidth}
\centering
\includegraphics[width=.48\linewidth]{fig5f.pdf}
\includegraphics[width=.48\linewidth]{fig5g.pdf}
\end{subfigure}
\label{fig}
\caption{\footnotesize Contributions to rate of adaptation from changes in trait variance and covariance. (a) $s$ varies with $N$ and $U$ being kept fixed ($N= 10^7$, $U=10^-5$). (b) $U$ varies with $N$ and $s$ being kept fixed ($N= 10^7$,$s=10^-2$). (c) $N$ varies with $s$ and $U$ being kept fixed ($s=10^-2$,$U=10^-5$).}
\end{figure}
% *****************************************************************************************************************
% ***********************************END*FIGURE*2******************************************************************
% *****************************************************************************************************************

@@

% *****************************************************************************************************************
% ***********************************FIGURE*3*COVARIANCE*TIME*SCALES*ARGUMENT*??***********************************
% *****************************************************************************************************************
\begin{figure}[h!]
\begin{subfigure}[b]{1\linewidth}
\label{a}
\caption{Two Dimensional Distribution \& Covariance Fluctations}
\includegraphics[width=0.48\linewidth]{fig3_N-10p09_c1-0d01_c2-0d01_U1-1x10pn5_U2-1x10pn5_exp1.pdf}
\includegraphics[width=0.48\linewidth]{fig9a.png}
\end{subfigure}
\begin{subfigure}[b]{1\linewidth}
\label{b}
\caption{Abundances Along Axis of Total Relative Fitness}
\includegraphics[width=1\linewidth]{fig9b.png}
\end{subfigure}
\label{Fig 5}
\caption{\footnotesize Evolution of the G-matrix over time. (a) Time dependent trajectories of trait one's variance and covariance are plotted with their mean values. Both are highly correlated in their behavior and fluctuate substantially over time. (b) Two dimensional trait distribution with two sets of classes identified. The first set, whose boxes are shaded in with patterned gray, represent the set of classes with 38 total beneficial mutations.  At time $t$, the population mean fitness lies on the line of constant fitness given by 38 total beneficial mutations; mean fitness is marked by the black dot. At time $t+tau_q$ the mean shifts to the next line of constant fitness associated with 39 total beneficial mutations. Here, $\tau_q \sim v^(-1)$ is the mean time between establishements described in \cite{desai2007beneficial}. (b) Depictions of the associated one dimensional traveling wave at times $t$ and $t+\tau_q$. Due to the Gaussian shape of the one dimensional traveling wave, the covariance between traits is set by proportions of classes within the peak. At time $t$ the group with 38 total beneficial mutations determine covariance, while at time $t+\tau_q$ it will be the group with 39 total beneficial mutations that sets covariance. In between these times the covariance is approximately a convex combination of its values at these two times.}
\end{figure}
% *****************************************************************************************************************
% ***********************************END*FIGURE*3******************************************************************
% *****************************************************************************************************************

Under beneficial mutation-selection balance, the time scale governing the dynamics of the G-matrix components is given by the mean time between establishements $\tau_q \sim 1/v(N,s,2U)$ described by \cite{desai2007beneficial}. Figure 3 shows the fluctuations in the variance of trait one, and its covariance with trait two.  The two are are highly correlated in their dynamics as a result of their adherence to equation 5. When the population is beneficial mutation-selection balance, the left hand side of this expression is approximately constant, and as a result, variance and covariance satisfy a linear relationship on average. This allows us to describe the dynamics of G by understanding the behavior of covariance over time, which is entirely driven by changes in the associated one dimensional fitness distribution depicted in figure 3c. The peak of the traveling wave represents the largest portion of the population. The frequencies of classes in this group, which are those whose fitness is equal to the population's mean fitness, to a large extent determine the covariance of the two dimensional distribution. After $\tau_q$ generations the mean fitness shift forward, and the peak of the one dimensional traveling wave will then coincide with next group of classes that lie on the next line of constant fitness, whose fitness subsequently equals the mean fitness, as shown in figure 3b. These major transitions in covariance are governed by the rate of adaptation. In between the times of these transitions, the covariance will simply be a convex combination of the two adjacent sets of classes that lie on the lines of constant fitness below and above the mean. Minor fluctuations in these dynamics are contributed by the residual portion of abundances not included in these two groups.\par

>>>>

\graphicspath{{figures/}}
\newcommand{\G}{\textbf{G}}

@@

For quantitative traits, genetic associations and their effects on adaptation are mathematically described by Lande's equation ($\Delta \bar{z} = \G \hspace{.05in}\beta$). The expression provides the expected change in the population mean of a trait $\Delta \bar{z}$ as the product of the selection gradient $\beta$ and what is known as the \G-matrix, whose components are the additive genetic covariances between traits. Lande’s result has been used to extend much of our understanding of how phenotypes evolve under selection, yielding a rich set of ideas that emphasize the role of the \G-matrix in directing adaptive evolution at both micro- and macro-evolutionary scales.\par
% * <masel@email.arizona.edu> 2017-07-29T12:43:16.916Z:
% 
% > When combined with the notions of adaptive landscapes
% The link to adaptive landscapes is not obvious, including to me. G-matrix theory is all continuous, and the only possible landscape is a phenotype-fitness map. The adaptive landscape you are going to deal with is a genotype-fitness map, ie a completely different object.
% 
% ^ <kgomez81@math.arizona.edu> 2017-08-07T14:10:30.176Z.

The behavior of the \G-matrix over time largely depends on what are the causes of the genetic correlations. Correlations between quantitative traits can arise for a number of reasons, but they often have a genetic basis \citep{Saltz2017}. They arise from either pleiotropy, linkage disequilibria, or are the product of trans acting elements; the first two that are thought to be the most significant factors. Historically, there have been two schools of thought what the principle causes of genetic correlations are. The first was put forth by the Edinburgh school, which held that pleiotropy was principle source, since linkage disequilibrium could rapidly decay by recombination \citep[Chapter~20]{fox2006evolutionary}. This belief was largely motivated by interests the relationships between parent-offspring phenotypes within randomly mating populations, and its implications for evolution. The alternative view came out of the Birmingham school, which largely experimented with inbred lines of plants. Geneticists in this group believed that linkage disequilibria was much more important; a fact that often reflected in their experiments. Presently, the Edinburgh view is more widely accepted, and as a result, most models for the evolution of the \G-matrix assume genetic correlations are pleiotropic and proceed with examining how evolutionary processes influence them. We know far less about the about the behavior genetic correlations from linkage disequilibria over time, but this is not much of an issue for most complex organisms.\par
% * <masel@email.arizona.edu> 2017-08-22T16:50:23.214Z:
% 
% >  most models for the evolution of the \G-matrix assume genetic correlations are pleiotropic and proceed with examining how evolutionary processes influence them
% An important exception to note is the "Bulmer effect", which I believe describes how selection can cause LD even in the absence of linkage, which then of course appears in G. This is distinct from our work where linkage to the background on which a mutation first appear is the cause of LD, albeit later shaped by selection. I believe there is some material on the Bulmer effect in Bruce Walsh's book. It is important to review it just enough to distinguish what you are doing from what Bulmer already did.
% 
% ^.
% * <masel@email.arizona.edu> 2017-08-22T16:47:43.713Z:
% 
% > the product of trans acting elements
% This is not self-explanatory, and insofar as I can guess what it means, it seems to me to be a special case of pleiotropy rather than a third distinct possibility.
% 
% ^.

@@

Lande’s equation describes the phenotypic evolution of populations with multiple characters from generation to generation.  Whether one can use this framework to uncover the history of natural selection in populations over large evolutionary timescales largely depends on the forces that influence the evolution of the \G-matrix.  Over the past few decades there has been a significant amount of effort to shed light on the matter of \G-matrix stability.  The analysis of genetic models for the adaptive evolution of multiple phenotypic traits indicate some of the conditions needed for stability\citep{Turelli1988,Jones2003}.  These works exclude the potential influences of linkage disequilibrium on the dynamics of variances and covariances over time, and as a result, our understanding of the role of linkage disequilibria in \G-matrix evolution remains unclear.  Nonetheless, the question is significant there is empirical evidence suggesting that in the \G-matrix is often conserved in natural populations \citep{Arnold1999,Roff2000,Steppan2002}.  However, much of what we know about \G-matrix stability corresponds the examining evolutionary forces on

@@

% *****************************************************************************************************************
% ***********************************FIGURE 1*2D*Travelling*Wave***************************************************
% *****************************************************************************************************************
\begin{figure}[h!]
\includegraphics[width=1\linewidth]{fig1a.pdf}
\label{Figure 1.}
\caption{\footnotesize Evolutuion of a two dimensional traveling wave. The distribution depicted was sampled from simulations.  Each box represents a class composed of individuals with an identical number of beneficial mutations in each trait. Classes with solid coloring form part of the bulk, while those with patterned interiors represent the stochastic front. The red dot marks the mean number of beneficial mutations in each trait. The dashed red line represents a lines of constant fitness. Classes whose centers lie along this line have relative fitness equal to the population mean, but vary in the number of beneficial mutations that have accumulated in each trait. Relative fitness increases along the perpendicular direction. Classes . Beneficial mutations and random sampling produce negligible effects in the changes of their abundances. In asexuals, only beneficial mutations that occur on the fittest backgrounds produce lineages that may eventually fix. These are marked with arrows into the stochastic front. Other lineages are out-competed and remain insignificant in size with respect to bulk. As new classes, the bulk expands in the direction of increasing relative fitness, and classes below the red line shrink and eventually disappear. As a result, the two dimensional trait distribution continually shifts in the direction of increasing relative fitness over time. (Simulation parameters: $N=10^9$, $s=10^{-2}$, and $U=10^{-5}$)}
\end{figure}
% *****************************************************************************************************************
% *******************************END*FIGURE 1**********************************************************************
% *****************************************************************************************************************

The dynamics of the bulk was modeled by a system of ordinary differential equations, in which abundances changed in proportion to their selective advantage \eqref{eq:1}. 
\begin{equation} \label{eq:1}
\dot{n}_{ij}(t) = (r_{ij}-\bar{r}) n_{ij}(t). 
\end{equation}
Abundances in the stochastic front were negligible in size relative to $N$ due to our parameter assumptions stated earlier.  Since the bulk represents the majority of population, we assumed that the frequencies of the bulk followed an analogous ODE.
\begin{equation} \label{eq:2}
\dot{p}_{ij}(t) = (r_{ij}-\bar{r}) p_{ij}(t) 
\end{equation}
These frequencies determine all of the quantitative genetic measures that are of interest to us.\par

@@

In order to investigate how of traits affect one another, we identified the contributions of each trait to the evolution of the population. The rate of adaptation can obtained by differentiating the population's mean relative fitness $\bar{r}=\sum_{ij} p_{ij} r_{ij}$, which is itself the sum of the two marginal means, $\bar{r_1}=\sum_{ij} p_{ij} is_1$ and $\bar{r_2}=\sum_{ij} p_{ij} j s_2$.  It follows that the total rate of adaptation is sum of rates of adaptation corresponding to each trait ($\dot{\bar{r}}=\dot{\bar{r_1}} +\dot{\bar{r_2}}$).  We can apply Lande's result to the two components $\dot{\bar{r_1}} $ and $\dot{\bar{r_2}} $ and rewrite them in terms of the trait variances $\sigma_1^2 = \sum_{ij} p_{ij} (r_i-\bar{r_1})^2$ and $\sigma_2^2 = \sum_{ij} p_{ij} (r_j-\bar{r_2})^2$, and their covariance $\sigma_{12} =\sum_{ij} p_{ij} (r_i-\bar{r_1})(r_j-\bar{r_2})$.  In doing so, we obtain the system expressed in \eqref{eq:5}.

@@

% Instead of hard-coding equation numbers, you can create auto-numbered links - ask Jason how. Right now the number doesn't match. You should also link this more clearly to the way it was presented in the introduction, ie you should explain that this is simply $\Delta \bar{z} = \G \hspace{.05in}\beta$ in the case of two traits both contributing directly to fitness

@@

The contributions of the stochastic front were incorporated into the adaptive evolution of the population by simulating each establishment event. We did this by identifying the first mutant lineage destined to establish among all mutants appearing in the stochastic front. Mutants awhich are parallel to the line shownwhich are parallel to the line shown appear according to a time-inhomogeneous Poisson process with rate functions given by the abundances of the adjacent classes from which they originated. Consequently, the arrival time $\tau_l$ of the $l^{th}$ mutant generated from the class $n_{ij}(t)$ having a mutation in trait $k$ can be used to form the random variable
% * <masel@email.arizona.edu> 2017-08-22T22:55:54.046Z:
% 
% > can be used to form the random variable
% This is confusing. The equation is for the cumulant for the inhomogeneous rate, the equation does not describe a random variable.
% 
% ^.
\[ 
\int_0^{\tau_l} U_k n_{ij}(t) dt ,
\] 
which has the distribution of $l$ i.i.d exponentially distributed summed random variables. By randomly sampling from the this distribution, we could then solve for each arrival time $\tau_l$. We then computed the probability of fixation using \eqref{eq:5}. The integral was numerically approximated up to an appropriate cutoff.  To determine if a mutation would establish, we sampled a uniform random variable and checked it against the calculated $\pi_{fix}$. The next class to establish was chosen accordingly and incorporated into the bulk with a random initial size generated with from \eqref{eq:4}.\par
% * <masel@email.arizona.edu> 2017-08-22T22:57:52.494Z:
% 
% >  \textbf{equation 5}
% Again, replace all these with auto-numbered links to specific equations, not to hard-numbered text.
% 
% ^.

@@

We allowed the simulation to run for 5,000 generations to allow the population to achieve beneficial mutation-selection balance before collecting data. This ensured that the transient effects from the initializing would not distort the statistics collected for the distribution. Following the burn period of the simulation, we recorded the existing classes in the bulk and their respective abundances. The data was used to compute the marginal mean fitness and variances of each trait, as well as their covariance. From these quantities we also computed the instantaneous rate of adaptation \eqref{eq:5}, the mean rate of fitness increase in each trait and the overall rate of adaptation.

@@

In each of our simulations we observed a 40\% decrease in the rate of adaptation of a trait as a result adding a second. The decrease was consistent across three parameter sets which varied either $N$, $s$ and $U$. These results also agree with predictions that can be derived from Desai and Fisher's theoretical work ($v(N,s,U)$). However, our simulations show that upon adding the second trait, the net change in $v_1$ arises from an increase in the first trait's variance and the formation of negative covariance. The first results in an increase in $v_1$, as a direct consequence of \eqref{eq:5}. This gain is more than offset by the negative covariance created between the two traits, whose magnitude is always substantially larger than the variance increase.  Figure 2 presents the calculated scaled mean variance of trait one, and its scaled covariance with the second trait, as a function of each varying population parameter. The graph demonstrates that within these parameter ranges, the new rate of adaptation $v_1$ as a percentage of the prior in every case is nearly 60\% less due solely to the large amounts of negative covariance that builds up between the two traits.\par

% *****************************************************************************************************************
% ***********************************FIGURE 2*Scaled Var/Cov Changes & Trait Reduction*****************************
% *****************************************************************************************************************
\begin{figure}[h!]
\begin{subfigure}[b]{1\linewidth}
\centering
\includegraphics[width=.48\linewidth]{fig5d.pdf}
\includegraphics[width=.48\linewidth]{fig5e.pdf}
\end{subfigure}
\begin{subfigure}[b]{1\linewidth}
\centering
\includegraphics[width=.48\linewidth]{fig5f.pdf}
\includegraphics[width=.48\linewidth]{fig5g.pdf}
\end{subfigure}
\label{fig}
\caption{\footnotesize Contributions to rate of adaptation from changes in trait variance and covariance. (a) $s$ varies with $N$ and $U$ being kept fixed ($N= 10^7$, $U=10^-5$). (b) $U$ varies with $N$ and $s$ being kept fixed ($N= 10^7$,$s=10^-2$). (c) $N$ varies with $s$ and $U$ being kept fixed ($s=10^-2$,$U=10^-5$).}
\end{figure}
% *****************************************************************************************************************
% ***********************************END*FIGURE*2******************************************************************
% *****************************************************************************************************************

@@

% *****************************************************************************************************************
% ***********************************FIGURE*3*COVARIANCE*TIME*SCALES*ARGUMENT*??***********************************
% *****************************************************************************************************************
\begin{figure}[h!]
\begin{subfigure}[b]{1\linewidth}
\label{a}
\caption{Two Dimensional Distribution \& Covariance Fluctations}
\includegraphics[width=0.48\linewidth]{fig3_N-10p09_c1-0d01_c2-0d01_U1-1x10pn5_U2-1x10pn5_exp1.pdf}
\includegraphics[width=0.48\linewidth]{fig9a.png}
\end{subfigure}
\begin{subfigure}[b]{1\linewidth}
\label{b}
\caption{Abundances Along Axis of Total Relative Fitness}
\includegraphics[width=1\linewidth]{fig9b.png}
\end{subfigure}
\label{Fig 5}
\caption{\footnotesize Evolution of the \G-matrix over time. (a) Time dependent trajectories of trait one's variance and covariance are plotted with their mean values. Both are highly correlated in their behavior and fluctuate substantially over time. (b) Two dimensional trait distribution with two sets of classes identified. The first set, whose boxes are shaded in with patterned gray, represent the set of classes with 38 total beneficial mutations.  At time $t$, the population mean fitness lies on the line of constant fitness given by 38 total beneficial mutations; mean fitness is marked by the black dot. At time $t+tau_q$ the mean shifts to the next line of constant fitness associated with 39 total beneficial mutations. Here, $\tau_q \sim v^(-1)$ is the mean time between establishements described in \cite{desai2007beneficial}. (b) Depictions of the associated one dimensional traveling wave at times $t$ and $t+\tau_q$. Due to the Gaussian shape of the one dimensional traveling wave, the covariance between traits is set by proportions of classes within the peak. At time $t$ the group with 38 total beneficial mutations determine covariance, while at time $t+\tau_q$ it will be the group with 39 total beneficial mutations that sets covariance. In between these times the covariance is approximately a convex combination of its values at these two times.}
\end{figure}
% *****************************************************************************************************************
% ***********************************END*FIGURE*3******************************************************************
% *****************************************************************************************************************

Under beneficial mutation-selection balance, the time scale governing the dynamics of the \G-matrix components is given by the mean time between establishements $\tau_q \sim 1/v(N,s,2U)$ described by \cite{desai2007beneficial}. Figure 3 shows the fluctuations in the variance of trait one, and its covariance with trait two. The two are are highly correlated in their dynamics as a result of their adherence to \eqref{eq:5}. When the population is beneficial mutation-selection balance, the left hand side of this expression is approximately constant, and as a result, variance and covariance satisfy a linear relationship on average. This allows us to describe the dynamics of \G-matrix by understanding the behavior of covariance over time, which is entirely driven by changes in the associated one dimensional fitness distribution depicted in figure 3c. The peak of the traveling wave represents the largest portion of the population. The frequencies of classes in this group, which are those whose fitness is equal to the population's mean fitness, to a large extent determine the covariance of the two dimensional distribution. After $\tau_q$ generations the mean fitness shift forward, and the peak of the one dimensional traveling wave will then coincide with next group of classes that lie on the next line of constant fitness, whose fitness subsequently equals the mean fitness, as shown in figure 3b. These major transitions in covariance are governed by the rate of adaptation. In between the times of these transitions, the covariance will simply be a convex combination of the two adjacent sets of classes that lie on the lines of constant fitness below and above the mean. Minor fluctuations in these dynamics are contributed by the residual portion of abundances not included in these two groups.\par
